---
title: "Homework 2"
author: "Tongtong Zhu"
date: "2022-10-04"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

### Read and clean the data 

```{r}
trans_ent = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

**Description of dataset**

The dataset `trans_ent` contains `r ncol(trans_ent)` columns and `r nrow(trans_ent)` rows. The variables include `line`, `station_name`, `station_latitude`, `station_longitude`, route (1-11), `entry`, `exit_only`, `vending`, `entrance_type`, `ada`. Data cleaning steps begin with data read, clean variable names, change the type of variable of Route columns (8-11) to character, arrange the sequence of columns, change the type of `entry` variable to logical. 

These data are not "tidy", because route number should be a variable. 

### Calculate the number of distinc stations

**Number of distinct stations**

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

**Number of distinct stations with ADA compliant**

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

**Proportion of station entrances / exits**

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

**Number of distinct stations serve A train, with ADA compliant**

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

**Summary**

There are total 465 distinct stations. 
There are 84 distinct stations with ADA compliant.
37.7% of station entrances/exits without vending allows entrance.
There are 60 distinct stations serve the A train. Of the stations that serve the A train, 17 are ADA compliant.


## Problem 2

### Read and clean Mr. Trash Wheel sheet

```{r}
mr_trash_wheel = 
  read_excel(
    "./data/Trash Wheel Collection Data.xlsx",
    sheet = 1,
    range = "A2:N550",
    skip = 1, 
    col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls, digits = 0)))
```

### Read and clean Professor Trash Wheel sheet

```{r}
pro_trash_wheel = 
  read_excel(
    "./data/Trash Wheel Collection Data.xlsx",
    sheet = 2,
    range = "A2:M97",
    skip = 1, 
    col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = NA,
    year = as.character(year)) %>% 
  relocate(sports_balls, .before = homes_powered)
```

### Add variable to both datasets and combine them

```{r}
mr_trash_wheel_new = mutate(mr_trash_wheel, wheel_type = "mr")
pro_trash_wheel_new = mutate(pro_trash_wheel, wheel_type = "pro")
com_data = bind_rows(mr_trash_wheel_new, pro_trash_wheel_new)

```

### Calcualte the total weight of trash & total number of sports balls

```{r}
sum(pro_trash_wheel$weight_tons)
```


```{r}
sum(mr_trash_wheel$sports_balls[which(mr_trash_wheel$year == "2020")])
```


**Description of new combining dataset**

The new dataset `com_data` contains `r nrow(com_data)` observations of `r ncol(com_data)` variables. The key variables include `dumpsters`, `year`, `weight_tons`, `volume_cubic_yards`, `plastic_bottles`, `polystyrene`, `cigarette_butts`, `glass_bottles`, `grocery_bags`, `chip-bags`, `sports_balls`, `homes_powered` and `wheel_type`. 

Compared with the original separate datasets `mr_trash_wheel` and `pro_trash_wheel`, the new combining dataset include a new column `wheel_type` to track the data from which trash wheel. `mr_trash_wheel` is identified as `wheel_type` "mr", and `pro_trash_wheel` is identified as `wheel_type` "pro". 

The total weight of trash collected by Professor Trash Wheel is `r sum(pro_trash_wheel$weight_tons)` tons. The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(mr_trash_wheel$sports_balls[which(mr_trash_wheel$year == "2020")])`.



## Problem 3

### Read and clean data in pols-month file

```{r}
pols_month = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = "-" ) %>% 
  mutate(
         president = if_else(prez_gop == 1, "gop", "dem"),
         year = as.integer(year)) %>% 
  arrange(year, month) %>% 
  mutate(month = month.name[as.integer(month)]) %>% 
  select(-prez_dem, -prez_gop, -day)

```

### Read and clean data in snp file

```{r}
snp = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(
    year = as.integer(year),
    year = if_else(year > 49, year + 1900, year + 2000),
    month = as.integer(month)) %>% 
  relocate(year,month) %>% 
  arrange(year, month) %>% 
  mutate(month = month.name[month]) %>% 
  select(-day) 

```

### Read and clean data in unemployment file

```{r}
unemployment = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_percent") %>% 
  mutate(month = month.name[match(month,month.abb)]) %>% 
  janitor::clean_names()

```


















